---
title: "Data Manipulation"
output: github_document
date: "2022-09-22"
---

```{r setup, echo=FALSE}
library(tidyverse)
```

## Load in the FAS Litters Data

```{r}
litters_df = read_csv("./data_import_examples/FAS_litters.csv")
litters_df = janitor::clean_names(litters_df)
```
## `Select`

Choose some columns and not others.

```{r}
select(litters_df, group, gd0_weight)
```
I can select a range of columns using `:`
```{r}
select(litters_df, group, gd0_weight:gd_of_birth)
```
Can use `-` to keep all variables EXCEPT a specified variable
```{r}
select(litters_df, -litter_number)
```

Use select to rename columns

```{r}
select(litters_df, GROUP = group, LITTer_NUmBer = litter_number)
```

There is a dedicated function in `dplyr` to rename variables that does the same thing.
```{r}
rename(litters_df, GROUP = group, LITTer_NUmBer = litter_number)
```

Select helpers enhance the `select()` function
```{r}
select(litters_df, starts_with("gd"))
```
Move litter_number to the first variable, keep every other variable
```{r}
select(litters_df, litter_number, everything())
```
Relocate does the same thing, relocates `litter_number` to the first column
```{r}
relocate(litters_df, litter_number)
```

Select is the key function! Does a lot of the things you would need, although other functions exist and do similar things.



