---
title: "Data Manipulation"
output: github_document
date: "2022-09-22"
---

```{r setup, echo=FALSE}
library(tidyverse)
```

## Load in the FAS Litters Data

```{r}
litters_df = read_csv("./data_import_examples/FAS_litters.csv")
litters_df = janitor::clean_names(litters_df)
```
## `Select`

Choose some columns and not others.

```{r}
select(litters_df, group, gd0_weight)
```
I can select a range of columns using `:`
```{r}
select(litters_df, group, gd0_weight:gd_of_birth)
```
Can use `-` to keep all variables EXCEPT a specified variable
```{r}
select(litters_df, -litter_number)
```

Use select to rename columns

```{r}
select(litters_df, GROUP = group, LITTer_NUmBer = litter_number)
```

There is a dedicated function in `dplyr` to rename variables that does the same thing.
```{r}
rename(litters_df, GROUP = group, LITTer_NUmBer = litter_number)
```

Select helpers enhance the `select()` function
```{r}
select(litters_df, starts_with("gd"))
```
Move litter_number to the first variable, keep every other variable
```{r}
select(litters_df, litter_number, everything())
```
Relocate does the same thing, relocates `litter_number` to the first column
```{r}
relocate(litters_df, litter_number)
```

Select is the key function! Does a lot of the things you would need, although other functions exist and do similar things.

## `Filter`

Removes rows or keeps rows that you intend to keep in your analysis.

We can filter observations for which `gd0_weight < 22`.

```{r}
filter(litters_df, gd0_weight < 22)
```

There are several logical operators we can use.

`==` is the test for equality, NOT a single equal sign, which is used to assign variables!

```{r}
filter(litters_df, gd_of_birth == 20)
```
Putting `!` outside logical operation gives you the inverse. Alternatively, you can use `!=`.
```{r}
filter(litters_df, !(gd_of_birth == 20))
filter(litters_df, gd_of_birth != 20)

```
Putting `!` outside logical operation is better for more complicated logical statements.
```{r}
filter(litters_df, gd0_weight >= 22, gd_of_birth == 20)
```
```{r}
filter(litters_df, group == "Mod8")
filter(litters_df, group %in% c("Con7", "Mod8")) #Group contains Con7 or Mod8
```
Real challenge in using filter is coming up with a way to specify the observations you want to keep!

## `Mutate`

Lets me modify an existing variable or create a new variable

```{r}
mutate(litters_df, wt_gain = gd18_weight - gd0_weight) #Create new variable wt_gain
```
```{r}
mutate(
  litters_df, 
  wt_gain = gd18_weight - gd0_weight,
  group = str_to_lower(group)) # Modifies all observations in group to lowercase
```
## `Arrange`

Puts observations in a specified order. Not as commonly used as others.

```{r}
arrange(litters_df, pups_born_alive) #In ascending order
```




